---
title: "GO, KEGG, GSEA analysis using PEL proteomics data"
output:
  html_document:
    df_print: paged
---

```{r}
library(dplyr)
library(tidyr)
library(data.table)
library(gdata)
library(reshape2)
library(ggplot2)
library(factoextra)
library(dendextend)
library(reshape2)
library(clusterProfiler)
library(org.Hs.eg.db)
# if necessary, we used the data similar to the data for heatmap and PCA analysis described in the previous section.
# set working directory
#setwd('C:/Users/SiSP/Desktop/Dr P/ChomPoo')
setwd('C:/Users/patipark/Dropbox/ChomPoo')
# get list of .csv files in the directory using "_table_Phi" pattern in the filename
TableAUC <- read.csv('Cell_GeneNames_BC3_BCBL_GTO_TY1.csv', header = TRUE, stringsAsFactors = FALSE)
# group by cell, select only rows with FBS treatment 
# this will be used as a denominator for each cell lines x treatments
# we divide each protein expression value with its own name from the control 'FBS' sample, cell-wise.
# the so-called 'normfun' function starts here
normfun <- function(D){
  T1 <- TableAUC %>% group_by(Cell)
  T2 <- TableAUC %>% group_by(Cell) %>% filter(Treatment %in% c('FBS'))
  TC <- T1 %>% select_('Cell') %>% unique
  
  A <- lapply(1:dim(TC)[1], function(x) {
    n <- TC$Cell[[x]]
    TM <- T1 %>% filter(Cell %in% n) %>% select_('Cell','Treatment')
    a <- T1[ grepl(n,T1$Cell),-c(1,2)]
    b <- T2[ grepl(n,T2$Cell),-c(1,2)]
    c <- apply(a,1,'/',b)
    
    d <- do.call(rbind,c)
    
    e <- cbind(as.data.frame(TM), as.data.frame(d)) 
    })
  matrix <- do.call(rbind,A)
  return(matrix)
} # end of the function
# we called the result as the matrix
matrix <- normfun(TableAUC)
# store the matrix as new variable with selected treaments
t3fill <- matrix %>% filter(Treatment %in% c('A4','BAY','TOFA'))
nt <- t3fill[,-c(1,2)]    # remove columns 1,2 that are Cell and Treatment names. Only numeric elements are required in PCA analysis.
# NA should be replaced by mean of available data using function 'f1'
f1 <- function(vec) {
  m <- mean(vec[is.finite(vec)])
  vec[!is.finite(vec)] <- m
  return(vec)
}
# make sure that resulting table is data frame
Ynt <- as.data.frame(apply(nt,2,f1)) # %>% round(7) if needed, use it to round number
########################################################
ZV <- data.frame('Cols' = which(apply(Ynt, 2, var) == 0)) # Identify zero variance columns
ZV$Genes <- rownames(ZV)  # make rownames as column 1
Ynt[,ZV$Cols] <- NULL     # remove the columns with zero variance
ZN <- colnames(Ynt)[colSums(is.na(Ynt)) > 0]  # Identify NA Inf columns
Ynt <- Ynt[,-which(names(Ynt) %in% ZN)] # remove the columns with NAs Infs
#head(Ynt)
#######################################################
```


PCA starts here
```{r}
pca1 = prcomp(Ynt, scale. = TRUE)
eigenvals <- pca1$sdev        # calculate sqrt of eigenvalues
loadfactor <- pca1$rotation   # calculate loadings factors
PCScores <- pca1$x   
# The PCScores matrix is converted to a data frame with appropriate row and column names
scores <- as.data.frame(PCScores)
# generate combined column vector from 'Cell' + 'Treatment'
t3fill.temp <- unite(t3fill,'Labels',c('Cell','Treatment'), remove = FALSE)
# assign as the vector as row names
rownames(scores) <- t3fill.temp$Labels
#head(scores)
```

We can add PCA analysis script below

Prepare data frame for heatmap
```{r}
# combine three tables by column together
Ynt.1 <- cbind(t3fill$Cell, t3fill$Treatment, Ynt)
colnames(Ynt.1)[c(1,2)] <- c('Cell','Treatment')
D <- melt(Ynt.1, id = c('Cell','Treatment')) 
# reshapt long table to wide table in which rows are rows are protein variables and columns are Cell_Treatment
Dd <- dcast(D,variable~Cell+Treatment)
# Select only rows without missing values and store as new 'mat' data frame
mat <- Dd[complete.cases(Dd),]

row.names(mat) <- mat$variable
mat2 <- mat[,-1]

library(RColorBrewer)
library(colorRamps)
breaksList = seq(-3, 3, by = 0.5)
cols <- colorRamps::matlab.like(13)
library(pheatmap)
set.seed(1)
hp <- pheatmap(mat2, 
               scale = 'row',                          # rescale by row
               clustering_method = 'ward.D2',          # method for clustering
               clustering_distance_rows = 'euclidean', # use euclidean distance
               clustering_distance_cols = 'euclidean', # use euclidean distance
               cluster_rows = TRUE,                    # boolean for cluster by rows
               cluster_cols = TRUE,                    # boolean for cluster by columns
               cellwidth = 10,                         # set cell width
               #cellheight = 10,                       # set cell height
               cutree_rows = 2,                        # set break gap between rows
               cutree_cols = 2,                        # set break gap between columns
               show_colnames = TRUE,                   # boolean for showing column names
               show_rownames = FALSE,                  # boolean for showing row names
               annotation_names_col = TRUE,            # boolean for showing column annotation
               fontsize = 5,
               color = cols,                           # color codes used in this heatmap
               breaks = breaksList)                    # range of each colors 

```

Extract data from heatmap's clustering function
```{r}
# create new table using data from pheatmap 
# get access to each cluster by $tree_row as the genes are in row dimension
mat2.order <- cbind(mat2[c(hp$tree_row[["order"]]),
                         hp$tree_col[["order"]]],
                    cluster = cutree(hp$tree_row, k = 2)[hp$tree_row[["order"]]])
A <- mat2.order   # assign to new variable for easier recall
```

Now, we will enrich KEGG pathway.
The scripts below provide fuctions necessary for GO terms splitting and mapping.
```{r}
# this section contains 3 functions
# 1.geneSplit
# 2.genemap
# 3.myEnrichKEGG
# this function is required to produce list of gene using chr [.] as a separator
geneSplit <- function(varin){
  temp <- lapply(1:length(varin), function(x) {
                                  f <- strsplit(as.character(varin[[x]]),'[.]')
                                  g <- do.call(cbind,f)})
  varout <- do.call(rbind,temp)
  return(varout)
} # end of function

# First we map the genes to ENTREZID
# function for mapping gene start here
genemap <- function(G){
  id <- bitr(G, fromType = 'SYMBOL', 
                toType = c('ENTREZID','ENSEMBL','ENSEMBLPROT','UNIPROT'), 
                OrgDb = 'org.Hs.eg.db', drop = TRUE)
  unique.id <- id %>% 
    select_('SYMBOL','ENTREZID','ENSEMBL','ENSEMBLPROT','UNIPROT') %>% 
    unique() %>% 
    droplevels()
  return(unique.id)
} # end of function

# now we use enrichKEGG function using ENTREZID
# the function starts here
myEnrichKEGG <- function(G){
  # search for column name 'ENTREZID
  if ('ENTREZID' %in% colnames(G)){
    Gin <- G$ENTREZID
  } else {
    cat('No ENTREZID in data')}
  ggo.kegg <- enrichKEGG(gene = Gin,
                         organism = 'hsa',
                         keyType = 'kegg',
                         pvalueCutoff = 0.25,
                         pAdjustMethod = 'fdr',
                         minGSSize = 10,
                         maxGSSize = 500,
                         qvalueCutoff = 0.05,
                         use_internal_data = FALSE)
  c <- ggo.kegg@result # extract results from the function
  c.sort <- c[order(-c$Count),] # reorder rows based on Count column
  # calculate Gene Ratio (Genes/TotalGenes)
  # from GeneRatio column, extract 'chr' before and after '/' as number while omit '/'
  gsize <- as.numeric(sub("/\\d+$", "", as.character(c.sort$GeneRatio)))
  gcsize <- as.numeric(sub("^\\d+/", "", as.character(c.sort$GeneRatio)))
  c.sort$GR = gsize/gcsize # calculate GeneRatio as number and place in a new column
  # create dotplot displaying first 20 pathways with highest counts
  ctk <- ggplot(c.sort, aes(x = GR, # GR column on x-axis
                                   y = reorder(Description,Count), # Description on y-axis and placed in order based on Count
                                   color = qvalue, # qvalue as color of dot
                                   size = Count)) + # Count as size of dot
    geom_point() +
    scale_colour_gradientn(colours = c('red', 'blue')) +
    labs(x = 'Gene Ratio',
         y = 'Enriched KEGG pathways',
         color = 'FDR',
         size = 'Count') +
    theme_bw()
  
  output <- list('result' = c.sort,'plot' = ctk)
  
  return(output)
}

#KEGG analysis per condition
#We extract genes from each treatment condition based on cell lines.
#In each treatment,
#1. determine significant of the expression (FC-to-control)
#1.1 Normal distribution uses t-test
#1.2 Non-normal distribution uses Wilcoxon rank sum

# G = data frame containing:
# rows = genes
# columns = treatments with their expression value of each gene
# The function will keep only treatment columns
getKEGG <- function(G){
  # reshape table
  df <- G[, -which(names(G) %in% c('cluster'))] # remove non-treatment-related columns
  List.cols <- colnames(df) # get a list of treatment
  df.l <- lapply(seq(1, ncol(df), by = 1), function(i) df[, i]) # split data frame by column
  # this function will do something
  # the function starts here
  B <- lapply(seq(1, ncol(df), by = 1), function(x) {
    b <- data.frame('Gene' = rownames(df), 'Z' = df.l[[x]]) 
    b$log2.Z <- log2(b$Z) # log transform z-scores to get normal distribution within column
    b$log2.Z[which(!is.finite(b$log2.Z))] <- 0
    # if log2Z is near 0 = little fold-change of treatment vs. control
    # if log2Z is away from 0 = large fold-chnage of treatment vs. control
    # H0 = fold-change is 0
    # we calculate pvalue; if p > 0.05, we do not reject H0
    # if p < 0.05, we reject H0 
    SD <- 2
    b$pval.log2.Z <- SD*pnorm(-abs(b$log2.Z)) # 2 S.D.
    # only when pval < 0.05, we assign:
    # up-regulated genes as 1
    # down-regulated genes as -1, the others as 0
    b$status <- ifelse(b$pval.log2.Z < 0.05 & b$log2.Z > 0, 1,
                       ifelse(b$pval.log2.Z < 0.05 & b$log2.Z < 0, -1, 0))
    # we split data into list according to status -1,0,1
    b.split <- split(b, with(b, interaction(b$status)), drop = TRUE)
    
    # this function will map genes, enrich KEGG and plot figure using myenrichKEGG
    # the function starts here
    # D = input data frame with single status
    # E = name of treatment
    myfun1 <- function(D,E){
      d <- D %>% select_('Gene') %>% 
        unique() %>% 
        droplevels() # select up-regulated genes
      d.split <- geneSplit(as.character(d$Gene)) # split gene names using char [.]
      d.gmap <- genemap(d.split) # map genes with ENTREZID, ENSEMBL, SYMBOL
      d.gmap <- d.gmap[duplicated(d.gmap$ENTREZID == FALSE),] # remove duplicated rows using ENTREZID
      d.kegg <- myEnrichKEGG(d.gmap)
      st <- D$status[1]
      d.kegg$result$status <- rep(st, dim(d.kegg$result)[1])
      tx <- E
      d.kegg$result$treatment <- rep(tx, dim(d.kegg$result)[1])
      return(d.kegg)
    }
    
    b.out <- lapply(seq(1, length(b), by = 1), function(n) myfun1(b,List.cols[[x]]))
    # b.out <- lapply(seq(1, length(b.split), by = 1), function(n) mfun1(b.split[[n]],List.cols[[x]]))
  })
  return(B)
}

getKEGG2 <- function(G){
  # reshape table
  df <- G[, -which(names(G) %in% c('cluster'))] # remove non-treatment-related columns
  df$Gene <- rownames(df)
  d <- df %>% select_('Gene') %>% 
    unique() %>% 
    droplevels() # select up-regulated genes
  d.split <- geneSplit(as.character(d$Gene)) # split gene names using char [.]
  d.gmap <- genemap(d.split) # map genes with ENTREZID, ENSEMBL, SYMBOL
  d.gmap <- d.gmap[duplicated(d.gmap$ENTREZID == FALSE),] # remove duplicated rows using ENTREZID
  
  d.kegg <- enrichKEGG(gene = d.gmap$ENTREZID,
                         organism = 'hsa',
                         keyType = 'kegg',
                         pvalueCutoff = 1,
                         pAdjustMethod = 'BH',
                         minGSSize = 10,
                         maxGSSize = 1000,
                         qvalueCutoff = 1,
                         use_internal_data = FALSE)
  d.out <- d.kegg@result
  gene.size <- as.numeric(sub("/\\d+$", "", as.character(d.out$GeneRatio)))
  bg.size <- as.numeric(sub("^\\d+/", "", as.character(d.out$GeneRatio)))
  d.out$GR <- gene.size/bg.size # calculate GeneRatio as number and place in a new column
  d.coladd <- data.frame('GR' = d.out$GR)
  List.col <- colnames(d.out)
  loc.add <- which(List.col == 'GeneRatio')
  loc.drop <- which(List.col == 'GR')
  d.out[,loc.drop] <- NULL
  d.out <- as.data.frame(append(d.out,d.coladd , after = loc.add))
  
  addSYMBOL(d.out)
}

# this funtion will convert ENTRIZID to SYMBOL or vice versa
# D = data table from enrichKEGG containing 'geneID' column 
# ie.convert '9999/3245/2345/098813' to 'AAYB/UBQA/EIF4A/RNASP' 
addSYMBOL <- function(D){
  List.col <- colnames(D)
  loc <- which(List.col == 'geneID')
  mi <- D[,loc]
  mt <- strsplit(as.character(mi),'[/]')
  sym.list <- lapply(1:length(mt), function(x) {
    m <- lapply(1:length(mt[[x]]), function(y) mt[[x]][[y]])
    ms <- do.call(cbind,m)
    mc <- bitr(ms, fromType = 'ENTREZID', toType = 'SYMBOL', OrgDb = 'org.Hs.eg.db')
    ml <- paste0(mc$SYMBOL, collapse = '/')
  })
  sym.col <- data.frame( 'geneSYMBOL' = do.call(rbind,sym.list))
  Dd <- as.data.frame(append(D, sym.col, after = loc))
  return(Dd)
}

```

Run function to get list of KEGG pathways
```{r message=FALSE, warning=FALSE}
######################
# just take whole gene list for enrichKEGG
B.results <- getKEGG2(A)

#head(B.results)
```

Screen for KEGG pathway names containing 'signaling pathway'
```{r}
# subset only signaling pathways
Sig <- B.results %>% filter(grepl('.*signaling pathway',Description)) %>% droplevels()
# create new column with shorter Description
library(tm) # NLP package
mi <- Sig$Description
mt <- removeWords(as.character(mi),'signaling pathway')
loc.add <- which(colnames(Sig) == 'Description')
Sig <- as.data.frame(append(Sig, data.frame('ShortDescription' = mt), after = loc.add))
# log2 transform p.adjust.value
mi <- -log2(Sig$p.adjust)
loc.add <- which(colnames(Sig) == 'p.adjust')
Sig <- as.data.frame(append(Sig, data.frame('log2.p.adjust' = mi), after = loc.add))
```

Create first doplot showing GeneRatio and p.adjust.value of enriched KEGG pathway.
Ranking by p.adjust.value from high to low.
```{r}
library(RColorBrewer)
CC <- c('#ffffb2','#fecc5c','#fd8d3c','#f03b20','#bd0026')
range.cc <- length(CC)
min.p <- min(Sig$log2.p.adjust)
max.p <- max(Sig$log2.p.adjust)
range.p <- range(max.p - min.p)/range.cc
breaksList <- seq(min.p,max.p, by = range.cc)

#Sig.sort <- Sig[order(Sig$p.adjust),]
ct_fig <- ggplot(Sig, aes(x = log2.p.adjust, # GR column on x-axis
                y = reorder(ShortDescription,log2.p.adjust), # Description on y-axis and placed in order based on Count
                color = log2.p.adjust, # qvalue as color of dot
                size = GR*100)) + # Count as size of dot
  geom_point(aes(fill = log2.p.adjust), colour = 'black',pch = 21) +
  geom_vline(xintercept = -log2(0.05), 
             #linetype = 'dotted', 
             color = 'red',
             size = 1) +
  #scale_x_reverse() +
  scale_fill_gradientn(colors = CC, breaks = breaksList, labels = breaksList)  +
  scale_x_continuous(breaks = scales::pretty_breaks(n = max.p+1)) +
  # scale_x_discrete(limits = c('BCBL_A4','TY1_A4','TY1_BAY',
  #                             'GTO_A4','GTO_BAY','BCBL_BAY',
  #                             'BC3_A4','BC3_BAY','BC3_TOFA',
  #                             'BCBL_TOFA','GTO_TOFA','TY1_TOFA')) +
  guides(fill = FALSE) +
  # guides(fill = guide_legend(title.theme = element_text(size = 10,
  #                                                       face = 'plain',
  #                                                       colour = 'black',
  #                                                       angle = 0,
  #                                                       vjust = 1))) +
  theme_bw() +
  theme(axis.text.x = element_text(face = 'plain', color = 'black', 
                                   size = 10, angle = 0)) +
  labs(x = expression(-log[2]~(Adj.p-value)),
       y = 'Enriched KEGG pathways',
       size = 'Gene Ratio (%)')

ct_fig
```

Significant enriched KEGG pathway may be use in further analysis for relevant pathways in a particular condition.

We rank again accoring to GeneRatio (the number of genes falling within a KEGG pathway).
```{r}
library(RColorBrewer)
CC <- c('#ffffb2','#fecc5c','#fd8d3c','#f03b20','#bd0026')
range.cc <- length(CC)
min.p <- min(Sig$log2.p.adjust)
max.p <- max(Sig$log2.p.adjust)
range.p <- range(max.p - min.p)/range.cc
breaksList <- seq(min.p,max.p, by = range.cc)

#Sig.sort <- Sig[order(Sig$p.adjust),]
ct_fig2 <- ggplot(Sig, aes(x = log2.p.adjust, # GR column on x-axis
                y = reorder(ShortDescription,GR), # Description on y-axis and placed in order based on Count
                color = log2.p.adjust, # qvalue as color of dot
                size = GR*100)) + # Count as size of dot
  geom_point(aes(fill = log2.p.adjust), colour = 'black',pch = 21) +
  geom_vline(xintercept = -log2(0.05),
             #linetype = 'dotted',
             color = 'red',
             size = 1) +
  #scale_x_reverse() +
  scale_fill_gradientn(colors = CC, breaks = breaksList, labels = breaksList)  +
  scale_x_continuous(breaks = scales::pretty_breaks(n = max.p+1)) +
  # scale_x_discrete(limits = c('BCBL_A4','TY1_A4','TY1_BAY',
  #                             'GTO_A4','GTO_BAY','BCBL_BAY',
  #                             'BC3_A4','BC3_BAY','BC3_TOFA',
  #                             'BCBL_TOFA','GTO_TOFA','TY1_TOFA')) +
  guides(fill = FALSE) +
  #guides(size = FALSE) +
  # guides(fill = guide_legend(title.theme = element_text(size = 10,
  #                                                       face = 'plain',
  #                                                       colour = 'black',
  #                                                       angle = 0,
  #                                                       vjust = 1))) +
  theme_bw() +
  theme(axis.text.x = element_text(face = 'plain', color = 'black', 
                                   size = 10, angle = 0)) +
  labs(x = expression(-log[2]~(Adj.p-value)),
       y = 'Enriched KEGG pathways',
       size = 'Gene Ratio (%)')

ct_fig2
```

Prepare data for gseKEGG analysis using the function provided below.
```{r}
# This function will create geneList suitable for pathview package
# according to GC Yu's blog post,
# we create geneList with head = KEGG_ID, value = expression level
# M = data frame with rows = genes, columns = treatment
myGeneList <- function(M){
  G <- M
  ColNames <- colnames(G)
  if ('cluster' %in% ColNames){
    df <- G[,!names(G) %in% c('cluster') == TRUE]
  } else {
    df <- G
  }# remove non-treatment-related columns
  # get a list of treatment
  List.cols <- colnames(df)
  # this function will do something
  # the function starts here
  B <- lapply(seq(1:dim(df)[2]), function(n) {
    h <- data.frame('SYMBOL' = rownames(df), 'value' = df[,List.cols[n]])
    rownames(h) <- rownames(df)
    # this function will truncate any letters after [.] in the gene names with multiple AA fractions
    # ie. aaaa.aaab.aabc -> aaaa
    dropReps <- function(varin){
      temp <- lapply(1:length(varin), function(x) {
        f <- gsub("\\..*","",as.character(varin[[x]]))})
      varout <- do.call(rbind,temp)
      return(varout)
    }
    # Now we will truncate long SYMBOL
    h$SYMBOL <- dropReps(h$SYMBOL)
    id <- genemap(h$SYMBOL) # map SYMBOL with hsa database
    List.id <- unique(id$ENTREZID) # some SYMBOL may return similar ENTREZID, we need unique ENTREZID
    # now we will use List.id as a provider of unique ENTREZID,
    # then, we select data (SYMBOL, ENTREZID) from 'id' gene mapping table
    sel <- id %>% filter(ENTREZID %in% List.id) %>% select_('SYMBOL','ENTREZID') %>% unique()
    # merge columns by rows based on similar SYMBOL
    # resulting in a data frame with unique sYMBOL, ENTREZID, protein expression value
    hh <- merge(sel,h, by = 'SYMBOL', all = TRUE)
    # if there are repeat ENTREZID rows, we calculate mean of protein expression value
    h3 <- setDT(hh)[, mean(value), by = c('SYMBOL','ENTREZID')]
    colnames(h3)[3] <- 'value'
    # in order to be used in gseKEGG function, data must be sorted in descending order (max -> min values)
    v <- log2(h3$value) # create a vector of protein expression values
    v[!is.finite(v)] <- 0
    names(v) <- h3$ENTREZID # add 'name' attribute to the vector
    vin <- sort(v,decreasing = TRUE) # sort vector
  })
}
```

Prepare gene+expression level suitable for the function.
```{r message=FALSE, warning=FALSE}
B <- myGeneList(mat2)
head(B[[1]])
```

Now we run tha gseKEGG analysis.
```{r message=FALSE, warning=FALSE}
kk2 <- lapply(1:length(B), function(n) gseKEGG(geneList = B[[n]],  # order ranked geneList
                                               organism = 'hsa',
                                               keyType = 'kegg', # 'kegg', 'ncbi-geneid','ncbi-proteinid', 'uniprot'
                                               exponent = 1, # weight of each step
                                               nPerm = 1000, # permutation numbers
                                               minGSSize = 10, # minimal size of each geneSet for analyzing
                                               maxGSSize = 1000, # maximal size of gene annotated for testing
                                               pvalueCutoff = 1, # pvalue cutoff
                                               pAdjustMethod = 'fdr', # pvalue adjustment method ( "holm", "hochberg", "hommel", "bonferroni", "BH", "BY", "fdr", "none")
                                               verbose = FALSE, # print message or not
                                               use_internal_data = FALSE, # logical, use KEGG.db or latest online KEGG data
                                               seed = FALSE)) # logical
```

We extract data frame containing KEGG pathways, enrichment scores etc.
```{r}
temp <- lapply(seq(1,length(kk2),by = 1), function(i) {kk2[[i]]@result %>% select_('NES')})
NES.cols <- do.call(cbind,temp)
colnames(NES.cols) <- colnames(mat2)

ID.cols <- kk2[[1]]@result %>% select_('ID','Description')

fin.cols <- cbind(ID.cols, NES.cols) # NES matrix
head(fin.cols)
```

Prepare data frame for pheatmap function
```{r}
mat3 <- fin.cols[,-c(1,2)] # prepare input matrix for pheatmap

library(RColorBrewer)
library(colorRamps)
breaksList = seq(-2, 2, by = 4/11)
cols <- rev(RColorBrewer::brewer.pal(11,'RdYlGn'))

library(pheatmap)
hp2 <- pheatmap(mat3, 
               scale = 'row',        # 'row' 'column' 'none'
               cluster_rows = TRUE,                    # boolean for cluster by rows
               cluster_cols = TRUE,                    # boolean for cluster by columns
               clustering_method = 'ward.D2',          # method for clustering
               clustering_distance_rows = 'euclidean', # use euclidean distance
               clustering_distance_cols = 'euclidean', # use euclidean distance
               cellwidth = 10,                         # set cell width
               #cellheight = 10,                       # set cell height
               #cutree_rows = 4,                        # set break gap between rows
               #cutree_cols = 5,                        # set break gap between columns
               show_colnames = TRUE,                   # boolean for showing column names
               show_rownames = TRUE,                  # boolean for showing row names
               annotation_names_col = TRUE,            # boolean for showing column annotation
               fontsize = 5,
               color = cols,                           # color codes used in this heatmap
               breaks = breaksList)
```

Select specific pathways and plot heatmap again.
```{r}
f.Sig <- fin.cols %>% filter(grepl('.*signaling pathway',Description)) %>% droplevels()
library(tm) # NLP package
mi <- f.Sig$Description
mt <- removeWords(as.character(mi),'signaling pathway')
loc.add <- which(colnames(Sig) == 'Description')
f.Sig <- as.data.frame(append(f.Sig, data.frame('ShortDescription' = mt), after = loc.add))
mat4 <- f.Sig[,-c(1:3)]
rownames(mat4) <- f.Sig$ShortDescription

library(RColorBrewer)
library(colorRamps)
breaksList = seq(-2, 2, by = 4/11)
cols <- rev(RColorBrewer::brewer.pal(11,'RdYlGn'))

library(pheatmap)
hp3 <- pheatmap(mat4, 
               scale = 'row',        # 'row' 'column' 'none'
               cluster_rows = TRUE,                    # boolean for cluster by rows
               cluster_cols = TRUE,                    # boolean for cluster by columns
               clustering_method = 'ward.D2',          # method for clustering
               clustering_distance_rows = 'euclidean', # use euclidean distance
               clustering_distance_cols = 'euclidean', # use euclidean distance
               cellwidth = 10,                         # set cell width
               #cellheight = 10,                       # set cell height
               #cutree_rows = 4,                        # set break gap between rows
               #cutree_cols = 5,                        # set break gap between columns
               show_colnames = TRUE,                   # boolean for showing column names
               show_rownames = TRUE,                  # boolean for showing row names
               annotation_names_col = TRUE,            # boolean for showing column annotation
               fontsize = 5,
               color = cols,                           # color codes used in this heatmap
               breaks = breaksList)
```

Now, we want to find similarity of each cell+drug pair.
Correlation is calculated and display as a correlation matrix.
Correlation matrix
```{r}
# NES-based correlation
# scale NES value across treatment conditions
scaled.mat <- apply(fin.cols[,-c(1:2)],2, scale)
# assign to new variable
mtx <- scaled.mat[,-c(1:2)]

corr <- cor(mtx, 
            #na.rm = FALSE,
            use = 'everything',
            method = 'spearman') # compute correlation matrix

# reorder cor mat
reorder_cormat <- function(cormat){
# Use correlation between variables as distance
dd <- as.dist((1-cormat)/2)
hc <- hclust(dd)
cormat <-cormat[hc$order, hc$order]
}
# use reorder_cormat function
cormat <- reorder_cormat(corr)

m.corr <- melt(cormat)

f_corr <- ggplot(m.corr, aes(Var1, Var2, fill = value)) +
  geom_tile(height = 0.8, width = 0.8) +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red",
                       midpoint = 0, limit = c(-1,1), space = "Lab", 
                       name = "Spearman\nCorrelation") +
  theme_minimal() +
  coord_equal() +
  labs(x = "",y = "",fill = "Corr") +
  theme(axis.text.x = element_text(size = 8, angle = 45, vjust = 1, hjust = 1, 
                                   margin = margin(-3,0,0,0)),
        axis.text.y = element_text(size = 8, margin = margin(0,-3,0,0)),
        panel.grid.major = element_blank())
f_corr
```

From correlation matrix, we found that cells trated with STATi were clustered together, suggesting their high similarity degree of protein expression profile.

Next step, we will measure distance between each treatment in a particular cell line.
```{r}
# Prepare a data frame
a <- TableAUC
nt <- a[,-c(1,2)]    # remove columns 1,2 that are Cell and Treatment names. Only numeric elements are required in PCA analysis.
# NA should be replaced by mean of available data using function 'f1'
f1 <- function(vec) {
  m <- mean(vec[is.finite(vec)])
  vec[!is.finite(vec)] <- m
  return(vec)
}
# make sure that resulting table is data frame
Ynt <- as.data.frame(apply(nt,2,f1)) # %>% round(7) if needed, use it to round number
########################################################
ZV <- data.frame('Cols' = which(apply(Ynt, 2, var) == 0)) # Identify zero variance columns
if (is.data.frame(ZV) && nrow(ZV) != 0){  
  ZV$Genes <- rownames(ZV)  # make rownames as column 1
  Ynt[,ZV$Cols] <- NULL     # remove the columns with zero variance
}
ZN <- colnames(Ynt)[colSums(is.na(Ynt)) > 0]  # Identify NA Inf columns
if (length(ZN) == 0) {
  Ynt <- Ynt[,-which(names(Ynt) %in% ZN)] # remove the columns with NAs Infs
} else {
  Ynt <- Ynt
}

#Ynt[1:5,1:5]

```

